% !TEX root = altosaar-2020-thesis.tex
\acknowledgements{
\lettrine[image=true,lines=3]{design/H}{undreds} of people have contributed to this thesis expedition. I am grateful for Acknowledgements, to reflect on the village it takes to train a Ph.D.

Shivaji Sondhi---thank you for your enduring support. Shivaji first answered my sophomoric email when I was searching for a summer internship in college, and years later took me under his wing after two advisers moved their labs out of Princeton during my first year of graduate school.

To David Blei, thank you for taking a chance on me. In our first research meeting I asked Dave what the word prior means, that I had been hearing it in the lab’s reading group for the past few months. I wondered out loud whether it might be important for me to learn about. Without missing a beat Dave suggested I pick up Bishop's textbook. The same spirit of non-judgmental openness and freedom has enabled me to carve my own path through graduate school, hewn from Dave’s support as a role model in thinking and writing clearly.

To Rajesh Ranganath: thank you for picking me up by the nape and pointing me in the right direction innumerable times and at personal cost; for believing in me when I did not, and gently showing me that I stand a chance despite the odds. As but one example, Rajesh had an unimaginably compassionate response to my errors when writing our first paper together, sleep-deprived. He interpreted my blunders as mere confounders in lieu of ignorance, and accurately inferred an underlying serious health condition. The subsequent sleep apnea diagnosis at a young age changed my outlook on life and research. I could at last retain and understand information without rote memorization, and my systolic blood pressure dropped a solid twenty. In the few counterfactuals where I redo grad school with enthusiastic consent, I get to re-meet Rajesh in them all---he fills the void of unambiguously altruistic, heroic characters in our lives. Thank you for modeling thinking of and doing for others (especially when they are not around), for indefatigable craic and curiosity under duress; for boldly asking why.

Thank you to my Columbia family: Aaron Schein, Keyon Vafa, Scott Linderman, Kriste Krstovski, Stephan Mandt, Stella Lianou, Kui Tang, Robert Florida, Dustin Tran, James McInerney, Wesley Tansey, Jackson Loper, Daisy Nguyen, Dawen Liang, Alp Kucukelbir, Victor Veitch, Liping Liu, Thibault Sellam, Adji Dieng, Ali Mehmani, Ghada Almashaqbeh, Yunhao Tang, Yixin Wang, Giannis Karamanolakis, Oscar Chang, Olivia Winn, Francisco J. Rodríguez Ruiz, Da Tang, Iñigo Urteaga, Ghazal Fazelnia, Christian Naesseth, Drim Stokhuijzen, and Gonzalo Esteban Mena.

To Kyle Cranmer, your generous enthusiasm and appetite for science is unmatched, and your warm welcome made the New York University stop in my thesis journey enriching and productive. To other colleagues and friends at NYU, thank you for the impromptu hangouts, whiteboards, and celebrations: I acknowledge Will Whitney, Nikita Nangia, Zach Martin, Raghav Singal, Michael Albergo, Min Jae Song, Mukund Sudarshan, David Brandfonbrener, Alfredo Canziani, Cinjon Resnick, Xintian Han, Mark Goldstein, Aahlad Manas Puli, and Ilya Kostrikov.

At Google and DeepMind, I am grateful to Eugene Brevdo and Andriy Mnih for mentoring me during unforgettable internships. Any hint of reliable experimentation and code used for this thesis is a direct reflection of what you taught me. And in the South Bay during my time at Google, I am grateful to Tiiu Jürvetson for being the most graceful, generous roommate a lowly intern could ask for.

Thank you to my collaborators Kexin Huang, Jingwei Zhang, Andrew James Mercer-Taylor, Ethan Benjamin, Rohan Bansal, Tony Paek, Anjishnu Kumar, Abhishek Bhatia, Drishan Arora, Shixiang Shane Gu, and Eamon Bell.

Cristina Alvarez, Christina DeLa Cruz, Sol and José at Big Sister Cleaners, Julio and Raoul at Rapid Park Industries, Robert Rafulowitz, Aimée Zambrana, Rayfield Gibson, Sharnice Ottley, and José Luis Ortiz: you were often the only (joyous!) human contact on many long days entering and exiting otherwise nameless buildings and streets.

Through Princeton I am grateful to have met many kindred spirits, including Jamal Williams, Sam Ritter, Gecia Bravo Hermsdorff, Lee Gunderson, Maxim Zaslavsky, Tom Clayton, Kevin Crowley, Siddharth Mishra-Sharma, Andrew Hein, Adam Pellegrini, Tim Treuer, Mark Ioffe, Vanessa Lehner, Richard Da, Kelvin Mei, DJ Strouse, Vladimir Kirilin, Angela Radulescu, Jon Berliner, Andrew Hartnett, Allison Chaney, Zack Dulberg, Alan Morningstar, and Tejal Bhamre. Our long walks around campus and the world are like fireworks; I am grateful for your stamina for conversating, gregarious rabbit holes, and your readiness to disagree for hours about weakly-held opinions.

To Princeton staff Josko Plazonic, Bill Wichser, Kate Brosowsky, Barbara Mooring, Jessica Heslin, Mariangela Lisanti, Jim Olsen, Bob Austin, Herman Verlinde---you help set a high-water mark for a supportive, collegial work environment. Any staff was willing to help at the slightest instigation, even when I was scared to ask. It makes a difference.

For making New York City feel like home I am indebted to Toby Shorin, Alessandra Poblador, Bharat Srikishan, Elisa Muyl, Bernard Itaka Lebrón, Jeffrey Seeley, Amy McDermott, Genevieve Fried, Luc Cary, Marina Krygin, Andra Mihali, Kariina Altosaar, Jessica Forde, Will Geary, Louise Contino, Alexandra Rone Lang, Tessa Bonduelle, Laurent Charlin, Clarissa Moliterno, Hiie Saumaa, Maja Rudolph, Gustav Kalm, James Shen, Sarth Calhoun, Ruth Tupe, Sophie Kleber, Becca Dittrich, and the forró family.

To Hanif Jetha, Kyle Saikaley, Maryse Thomas, Bohdan Kulchytskyy, Michael King, Fredrik Folkestad, Briana Salas, Francis Megna, Colin Raffel, Sarah Hagi, Christine Li, Ben Poole, Patrick Murumi Njoroge, Raul Altosaar, Laura Cooper Hall, David Dohan, Jasmine Collins, Dylan Cable, David Aasen, Jakob Foerster, Sergey Bartunov, Tyeler Matsuo, Eva Altosaar, Natasha Jaques, Vanessa Chazelle, Faris Haddad, Aleksi Altosaar, Courtney Yusuf, Rebecca Choong-Wilkins, Nick Greaves-Tunnell, Janine, Peter, and Justin Cary, Georgie Rubens, Solomon Sonya, and Cathy Ji---thank you for reminding me from near and far about what is important in life.

To my parents: Illimar, for always picking up the phone and infectious enthusiasm for the world; Tiiu, for assuaging rational and irrational panics in ways only a mother can.

Finally, I recognize Caroline Martin: if this is but a crisis with you, I would rejoice in submitting to considerably greater mid-pandemic quarantines, job hunts, and theses.
\sectionline
}